```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sbm)
library(ggplot2)
library(igraph)
library(knitr)
library(alluvial)

```

## Data Preparation and pre-processing
```{r}
lnc <- read.table("~/CLASS/Justin/data/lncRNA_interaction.txt", "\t",header = T,finc = T)

#only RNA-protein interactions
rna_protein <- lnc[lnc$level == "RNA-Protein",]

#only lncRNA-protein interactions
lnc_protein <- rna_protein[rna_protein$ncType == "lncRNA" & rna_protein$tarType == "protein",]

#only human lncRNA-protein interactions
human_lnc_protein <- rna_protein[rna_protein$ncType == "lncRNA" & rna_protein$tarType == "protein"& rna_protein$organism == "Homo sapiens",]

#filter out other information for adjacency matrix
human_interaction <- dplyr::select(human_lnc_protein, ncName, tarName)
rownames(human_interaction)<-NULL

human_interaction$ncName <- factor(human_interaction$ncName)
human_interaction$tarName <- factor(human_interaction$tarName)

f_lnc <- read.table("~/CLASS/Justin/data/filtered_lncRNA_Protein.txt", ",",header = F)

f_lnc$V1 <- factor(f_lnc$V1)
f_lnc$V2 <- factor(f_lnc$V2)

#filter out degrees <= 1
f_human_interaction <- human_interaction[human_interaction$ncName %in% f_lnc$V1 & human_interaction$tarName %in% f_lnc$V2,]

f_human_interaction$ncName <- factor(f_human_interaction$ncName)
f_human_interaction$tarName <- factor(f_human_interaction$tarName)

#convert the filtered lncRNA-protein interactions to an adjacency matrix
tmp_matrix <- table(f_human_interaction)

tmp <- data.frame(rbind(tmp_matrix))

tmp <- as.matrix(tmp)

#number of lncRNA 9630
num_of_lnc <- nrow(tmp)
#number of protein 290
num_of_protein <- ncol(tmp)

write.csv(f_human_interaction, file = '~/CLASS/Justin/CSCI3352-Project/01_Data_Preparation/results/Human_Lnc_Protein_interaction_Meta_Data.csv')

```

## LPBNI
```{r}
B = t(t(tmp))
nc = nrow(B)
tar = ncol(B)
W=matrix(0,tar,tar)
deci=matrix(0,tar,nc)

for (i in 1:tar)
{
gmq=matrix(rep(B[ ,i],tar),nc,tar)*B
q=apply(gmq,2,function(gmq,a) gmq/a,a=matrix(rowSums(B),nc,1))
W[i,1:tar]=(1/colSums(B))*colSums(q)        
}

for (j in 1:nc)
{
f0=B[j, ]
f1=W %*% f0
deci[1:tar,j]=f1
}
```


## Bipartite SBM
```{r}
# Plot the adjacency matrix
plotMyMatrix(tmp, dimLabels = list(row = 'nc', col = 'tar') )

# Estimate bipartite sbm of our adjacency matrix
BipartiteSBM <- 
  tmp %>% 
  estimateBipartiteSBM(model = 'bernounci', estimOptions = list(verbosity = 0, plot = FALSE, nbCores = 8))

plot(BipartiteSBM, dimLabels = list(row = 'nc', col = 'tar'))

BipartiteSBM$storedModels %>% kable()

#when nbparams = 257, 270, 283 and nblocks = 31, 32, 33 we have the lowest ICL
BipartiteSBM$storedModels %>% 
  ggplot() + aes(x = nbBlocks, y = ICL) + geom_line() + geom_point(alpha = 0.5)
ggsave('~/CLASS/Justin/CSCI3352-Project/01_Data_Preparation/figures/ICL_trend_No_filtered.png')

#set models with 33 clusters, 12 lncRNA clusters, 21 protein clusters
BipartiteSBM$setModel(33)

BipartiteSBM$print()

BipartiteSBM$blockProp

BipartiteSBM$probMemberships

lnc_clust <- BipartiteSBM$memberships

#overview of row clusters labels
table(unlist(lnc_clust$row))
#overview of column clusters labels
table(unlist(lnc_clust$col))

#retrieve which proteins are labeled cluster-20
ind <- which(lnc_clust$col == 20)

pro_list <- colnames(tmp)
lnc_list <- rownames(tmp)

clust_20 <- pro_list[which(lnc_clust$col == 20)]
